Liveness-Based Garbage Collection for Lazy Languages

### 2015-06-05 paper/paper.tex
the statement of undecidability

The  solution  is  to   also  treat  the  operations  mentioned  above
symbolically.  We  introduce three  new  symbols.   First, \bcar\  and
\bcdr\ are defined as:
\begin{align*}
 \bcar\sigma &\triangleq \{\alpha \mid 0\alpha \in \sigma\}\\
 \bcdr\sigma &\triangleq \{\alpha \mid 1\alpha \in \sigma\}
\end{align*}
and \cbar\ as:
\begin{align*}
 \cbar\sigma \triangleq\; & \emptyset~\mbox{if}~\sigma = \emptyset\\
                       & \{\epsilon\} ~\mbox{otherwise}
\end{align*}
Note that \cbar\ is idempotent. 
We can now rewrite the \CONS\ and the \CAR\ rules of $\mathit{ref}$ as:
\begin{align*}
&\mathit{ref\/}((\CONS~x~y),\sigma,\Lfonly)
= x.\bcar\sigma \cup y.\bcdr\sigma  \label{eqn:mod-cons},~
\mbox{and} \\
&\mathit{ref\/}((\CAR~x),\sigma,\Lfonly)
          =   x.\cbar\sigma \cup x.0\sigma
\end{align*}
and the \Lfunonly\ rule
for \SIF\ as:
\begin{align*}
\mathcal{L}((\SIF~x~e_1~e_2),\sigma,\Lfonly) =
                    &\mathcal{L}(e_1,\sigma,\Lfonly)~\uplus
        \mathcal{L}(e_2,\sigma,\Lfonly)~\uplus\\
        &[\,\mathit{ep} \mapsto  x.\cbar \sigma]
\end{align*}
The rules for \CDR, \PRIM\ and  \NULLQ\  are also modified in a manner
similar to \CAR.
The  defintions of \bcar\, \bcdr\ and \cbar\  induce   the  following
relation $\hookrightarrow$ over sets of access paths:
\begin{align*}
  &\sigma_1\bcar\sigma_2  \hookrightarrow
  \sigma_1\sigma_2'\mbox{ where } \sigma_2 = \acar\sigma_2',\\
&\sigma_1\bcdr\sigma_2  \hookrightarrow
  \sigma_1\sigma_2'\mbox{ where } \sigma_2 =
  \acdr\sigma_2',~\mbox{and}\\
&\sigma_1\cbar \sigma_2 \hookrightarrow \epsilon, \mbox{ if } \sigma \ne
  \emptyset \\
&\sigma_1\cbar \emptyset \hookrightarrow \emptyset
\end{align*}
The reflexive  transitive closure of  $\hookrightarrow$ will
be   denoted    as   $\stackrel{*}{\hookrightarrow}$.    The
following proposition relates  the original and the modified
liveness rules.

@@

%% We could convert the rules defining $\hookrightarrow$ into productions
%% and  add  them  to the  grammar.  However,  this  results in  an  {\em
%%   unrestricted   grammar}~\cite{hopcraft90toc},  and   the  membership
%% problem for  such grammars is undecidable. \comment{Rewrite  it to say
%%   that this is a CFG with  3 unrestricted rules and the membership for
%%   even this grammar  is undecidable} \comment{State the undecidability
%%   of the decision problem above.}

  We circumvent the problem by over-approximating the CFG generated by
  the  analysis  to  {\em  strongly  regular}  CFGs  which  have  easy
  translations to  non-deterministic finite state  automata (NFA). The
  NFAs are then simplified on the basis of the $\hookrightarrow$ rules
  to  enable  checking of  membership  of  forward  access paths.  The
  resulting NFAs are finally converted  to DFAs for use during garbage
  collection.  \comment{Fix all comments and produce first draft}
%------------------------------------------------------------%
\subsubsection{Approximating CFGs using NFAs.}
\label{sec:NFA-approx}
\comment{Change the language and the example}
We use the
algorithm  by  Mohri  and Nederhof~\cite{mohri00regular}   to
approximate a CFG  to a   {\em strongly regular\/} grammar.  The
transformation has  the  property that  if $\Lv$  is a
non-terminal  in the grammar  $G$ and  $G'$ is  the grammar  after the
Mohri-Nederhof transformation, then  ${\mathscr L}_G(\Lv) \subseteq
{\mathscr
  L}_{G'}(\Lv)$.    This  is   required  for the
approximation to be safe with respect to liveness.

We show the \var{\Lanv{9}{\py}} grammar after the Mohri-Nederhof
transformation. We  pick the only production that  is affected by
the transformation---the production for \Df{\append}{1}.  The
production for \Uf{\append}{1}, while  recursive, is already in
strongly  regular  form  and   is  therefore  unaffected  by  the
transformation.
 \begin{eqnarray*}
  \var{\Df{\append}{1}}
  &\rightarrow&               \acar\bcar \var{\Df{\append}{1}}'  \mid
  \acdr\var{\Df{\append}{1}} \\
\var{\Df{\append}{1}}'  &\rightarrow&   \bcdr\var{\Df{\append}{1}}'
\mid \epsilon
\end{eqnarray*}
The  languages
generated for  \var{\Df{\append}{1}} in  the original grammar  and the
new       grammar      are       \acdr$^i$\acar\bcar\bcdr$^i$      and
\acdr$^*$\acar\bcar\bcdr$^*$, showing a  loss of  precision.

\subsubsection{Removing the $\cbar$ edges}
Once the NFA's are created, we need to remove the transitions on
$\cbar$. This edge is replaced by an $\epsilon$ transition and the
$\cbar$ edge is deleted. If the original grammar had a non-null
$\sigma$ then the NFA reaches a final state throught this path
otherwise it does not reach a final state.\comment{Need to rewrite
this}
%------------------------------------------------------------%
\subsubsection{Transforming NFAs to Accept Forward Paths:}
%\label{sec:NFA-elim}
The strongly regular CFGs obtained after the Mohri-Nederhof
transformation are
  first converted  into  NFAs.  The        algorithm         described
      in
Figure~\ref{algo:simplify-NFA} converts  an NFA \ntbar\  to a NFA
\nt\ such that  $\lang{\ntbar} \stackrel{*}{\hookrightarrow}
\lang{\nt}$, where {\nt} accepts forward paths only.
\comment{Rewrite the algorithm to handle \cbar and update the example
automata}
Thus  \nt\  can  be  used  to  check
membership of forward paths.

\begin{figure*}[t]
\begin{boxedminipage}{\textwidth}
\begin{center}
  \raggedright  {\bf  Input:}  NFA \ntbar\  with  underlying
  alphabet $\lbrace \acar, \acdr, \bcar, \bcdr\rbrace$ \\
%
  \raggedright{\bf Output:} NFA  \nt\ with  underlying
  alphabet $\lbrace \acar, \acdr \rbrace$ such that
  $\lang{\ntbar} \stackrel{*}{\hookrightarrow} \lang{\nt}$.
  \\
%

  \raggedright{\bf Steps:}
 %% \begin{algorithmic}
%%    \COMMENT Fix the algorithm issue
%%    %% \STATE $i \leftarrow 0$
%%    %% \STATE $\nt_0 \leftarrow$ Equivalent NFA of \ntbar\
%%    %% without $\epsilon$-moves \cite{hopcraft90toc}

%%    %% \REPEAT
%%    %% \STATE $\nt'_{i+1} \leftarrow \nt_i$
%%    %% \FORALL{states $q$ in $\nt_i$ such that $q$ has an incoming
%%    %%   edge from $q'$ with label $\bcar$ and
%%    %%   outgoing edge to $q''$ with label $\acar$}
%%    %% \STATE add an edge in $\nt'_{i+1}$ from $q'$ to $q''$ with
%%    %% label $\epsilon$.
%%    %% \COMMENT{bypass  $\bcar\acar$ using $\epsilon$}
%%    %% \ENDFOR

%%    %% \FORALL{states $q$ in $\nt_i$ such that $q$ has an incoming
%%    %%   edge from $q'$ with label $\bcdr$ and outgoing edge to
%%    %%   $q''$ with label $\acdr$}
%%    %% \STATE add an edge in $\nt'_{i+1}$ from $q'$ to $q''$ with
%%    %%   label $\epsilon$.
%%    %% \COMMENT{bypass  $\bcdr\acdr$ using $\epsilon$}
%%    %% \ENDFOR

%%    %% \STATE   $\nt_{i+1}    \leftarrow$   Equivalent   NFA   of
%%    %% $\nt'_{i+1}$ without $\epsilon$-moves

%%    %% \STATE $i \leftarrow i+1$
%%    %% \UNTIL ($\nt_{i} = \nt_{i-1}$)

%%    %% \STATE $\nt \leftarrow \nt_i$
%% %   \STATE delete all edges with label \bcar\ or \bcdr\ in \nt.
%%  \end{algorithmic}
\end{center}
\end{boxedminipage}
 \caption{Algorithm for transforming an NFA to  accept forward
   paths only.}\label{algo:simplify-NFA} \figrule
\end{figure*}

>>>>

The  solution  is  to   also  treat  the  operations  mentioned  above
symbolically.  We  introduce three  new  symbols.   First, \bcar\  and
\bcdr\ are defined as:
\begin{align*}
 \bcar\sigma &\triangleq \{\alpha \mid 0\alpha \in \sigma\}\\
 \bcdr\sigma &\triangleq \{\alpha \mid 1\alpha \in \sigma\}
\end{align*}
and \cbar\ as:
\begin{align*}
 \cbar\sigma \triangleq\; & \emptyset~\mbox{if}~\sigma = \emptyset\\
                       & \{\epsilon\} ~\mbox{otherwise}
\end{align*}
Note that \cbar\ is idempotent. 
We can now rewrite the \CONS\ and the \CAR\ rules of $\mathit{ref}$ as:
\begin{align*}
&\mathit{ref\/}((\CONS~x~y),\sigma,\Lfonly)
= x.\bcar\sigma \cup y.\bcdr\sigma  \label{eqn:mod-cons},~
\mbox{and} \\
&\mathit{ref\/}((\CAR~x),\sigma,\Lfonly)
          =   x.\cbar\sigma \cup x.0\sigma
\end{align*}
and the \Lfunonly\ rule
for \SIF\ as:
\begin{align*}
\mathcal{L}((\SIF~x~e_1~e_2),\sigma,\Lfonly) =
                    &\mathcal{L}(e_1,\sigma,\Lfonly)~\uplus
        \mathcal{L}(e_2,\sigma,\Lfonly)~\uplus\\
        &[\,\mathit{ep} \mapsto  x.\cbar \sigma]
\end{align*}
The rules for \CDR, \PRIM\ and  \NULLQ\  are also modified in a manner
similar to \CAR.
The  defintions of \bcar\, \bcdr\ and \cbar\  induce   the  following
relation $\hookrightarrow$ over sets of access paths:
\begin{align*}
  &\sigma_1\bcar\sigma_2  \hookrightarrow
  \sigma_1\sigma_2'\mbox{ where } \sigma_2 = \acar\sigma_2',\\
&\sigma_1\bcdr\sigma_2  \hookrightarrow
  \sigma_1\sigma_2'\mbox{ where } \sigma_2 =
  \acdr\sigma_2',~\mbox{and}\\
&\sigma_1\cbar \sigma_2 \hookrightarrow \epsilon, \mbox{ if } \sigma \ne
  \emptyset 
\end{align*}
The reflexive  transitive closure of  $\hookrightarrow$ will
be   denoted    as   $\stackrel{*}{\hookrightarrow}$.    The
following proposition relates  the original and the modified
liveness rules.

@@

We circumvent  the problem by over-approximating the  CFG  by
non-deterministic finite state automata (NFA). The NFA are then
simplified using the $\hookrightarrow$ rules. Finally the simplified
NFAs are converted to DFAs.

%------------------------------------------------------------%
\subsubsection{Approximating CFGs using NFAs.}
\label{sec:NFA-approx}
 We use  the algorithm by Mohri  and Nederhof~\cite{mohri00regular} to
 approximate   a  CFG   to   a  {\em   strongly  regular\/}   grammar.
 \comment{Rewrite  this after  constructing the  example: We  show the
   grammar for ????? after  the Mohri-Nederhof transformation. We pick
   the only  production that  is affected by  the transformation---the
   production    for     \Df{\append}{1}.     The    production    for
   \Uf{\append}{1},  while recursive, is  already in  strongly regular
   form and is therefore unaffected by the transformation.
 \begin{eqnarray*}
  \var{\Df{\append}{1}}
  &\rightarrow&               \acar\bcar \var{\Df{\append}{1}}'  \mid
  \acdr\var{\Df{\append}{1}} \\
\var{\Df{\append}{1}}'  &\rightarrow&   \bcdr\var{\Df{\append}{1}}'
\mid \epsilon
\end{eqnarray*}
The  languages
generated for  \var{\Df{\append}{1}} in  the original grammar  and the
new       grammar      are       \acdr$^i$\acar\bcar\bcdr$^i$      and
\acdr$^*$\acar\bcar\bcdr$^*$, showing a  loss of  precision.

Once the NFA's are created, we need to remove the transitions on
$\cbar$. This edge is replaced by an $\epsilon$ transition and the
$\cbar$ edge is deleted. If the original grammar had a non-null
$\sigma$ then the NFA reaches a final state throught this path
otherwise it does not reach a final state.}

%------------------------------------------------------------%
\subsubsection{Transforming NFAs to Accept Forward Paths:}
%\label{sec:NFA-elim}
The   strongly  regular   CFGs  obtained   after   the  Mohri-Nederhof
transformation are first converted into NFAs.  The algorithm described
in  Figure~\ref{algo:simplify-NFA} converts  an NFA  \ntbar\ to  a NFA
\nt\    such    that   $\lang{\ntbar}    \stackrel{*}{\hookrightarrow}
\lang{\nt}$, where {\nt} accepts forward paths only.  \comment{Rewrite
  the algorithm to handle \cbar  and update the example automata} Thus
\nt\ can be used to check membership of forward paths.

\begin{figure*}[t]
\begin{boxedminipage}{.5\textwidth}
\begin{center}
  \raggedright  {\bf  Input:}  NFA \ntbar\  with  underlying
  alphabet $\lbrace \acar, \acdr, \bcar, \bcdr\rbrace$ \\
%
  \raggedright{\bf Output:} NFA  \nt\ with  underlying
  alphabet $\lbrace \acar, \acdr \rbrace$ such that
  $\lang{\ntbar} \stackrel{*}{\hookrightarrow} \lang{\nt}$.
  \\

  \raggedright{\bf Steps:}
 \begin{algorithmic}
   \STATE $i \leftarrow 0$
   \STATE $\nt_0 \leftarrow$ Equivalent NFA of \ntbar\
   without $\epsilon$-moves \cite{hopcraft90toc}

   \REPEAT
   \STATE $\nt'_{i+1} \leftarrow \nt_i$
   \FORALL{states $q$ in $\nt_i$ such that $q$ has an incoming
     edge from $q'$ with label $\bcar$ and
     outgoing edge to $q''$ with label $\acar$}
   \STATE add an edge in $\nt'_{i+1}$ from $q'$ to $q''$ with
   label $\epsilon$.
   \COMMENT{bypass  $\bcar\acar$ using $\epsilon$}
   \ENDFOR

   \FORALL{states $q$ in $\nt_i$ such that $q$ has an incoming
     edge from $q'$ with label $\bcdr$ and outgoing edge to
     $q''$ with label $\acdr$}
   \STATE add an edge in $\nt'_{i+1}$ from $q'$ to $q''$ with
     label $\epsilon$.
   \COMMENT{bypass  $\bcdr\acdr$ using $\epsilon$}
   \ENDFOR

   \STATE   $\nt_{i+1}    \leftarrow$   Equivalent   NFA   of
   $\nt'_{i+1}$ without $\epsilon$-moves

   \STATE $i \leftarrow i+1$
   \UNTIL ($\nt_{i} = \nt_{i-1}$)

   \STATE $\nt \leftarrow \nt_i$
  \STATE delete all edges with label \bcar\ or \bcdr\ in \nt.
 \end{algorithmic}
\end{center}
\end{boxedminipage}
\begin{boxedminipage}{.5\textwidth}
The algorithm repeatedly introduces  $\epsilon$ edges to bypass a pair
of consecutive edges labelled  \bcar\acar\ or \bcdr\acdr.  The process
is continued till  a fixed point is reached.  When  the fixed point is
reached,   the  resulting   NFA  contains   all   possible  reductions
corresponding to all the paths in the original NFA.  The proofs of the
termination and correctness  of th The algorithm repeatedly introduces  $\epsilon$ edges to bypass a pair
of consecutive edges labelled  \bcar\acar\ or \bcdr\acdr.  The process
is continued till  a fixed point is reached.  When  the fixed point is
reached,   the  resulting   NFA  contains   all   possible  reductions
corresponding to all the paths in the original NFA.  The proofs of the
termination and correctness  of th The algorithm repeatedly introduces  $\epsilon$ edges to bypass a pair
of consecutive edges labelled  \bcar\acar\ or \bcdr\acdr.  The process
is continued till  a fixed point is reached.  When  the fixed point is
reached,   the  resulting   NFA  contains   all   possible  reductions
corresponding to all the paths in the original NFA.  The proofs of the
termination and correctness  of th The algorithm repeatedly introduces  $\epsilon$ edges to bypass a pair
of consecutive edges labelled  \bcar\acar\ or \bcdr\acdr.  The process
is continued till  a fixed point is reached.  When  the fixed point is
reached,   the  resulting   NFA  contains   all   possible  reductions
corresponding to all the paths in the original NFA.  The proofs of the
termination and correctness  of th

\end{boxedminipage}
 \caption{Algorithm for transforming an NFA to  accept forward
   paths only.}\label{algo:simplify-NFA} \figrule
\end{figure*}
